Cropland percentage bias and accuracy, part II

### 2017-01-27 Rmd/c10-location-impact.Rmd
Edits to manuscript to 27 Jan, 2017. Added/edited code in c3-cropland-1.Rmd (saved aggregated cropland and area weights raster lists to output rda files), c4-cropland-2.Rmd (calculated overall cropland areas by dataset and with different scales of aggregation).

# yield and carbon data
load("external/ext_data/yieldprod-1km.rda")
load("external/ext_data/carbon-tier1.rda")

@@

### Identify high value areas
```{r, eval = FALSE}
# top decile of...
# ...yield
yclass <- lapply(cyld_agg$f1, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

@@

mudists <- lapply(list(yclass, pclass, cclass, effclass), function(x) {
  # x <- yclass
  mudist <- sapply(2:5, function(y) {  # y <- 2
    print(paste0("...", y))
    a <- rasterToPoints(x$gti, fun = function(x) x == 1)
    b <- rasterToPoints(x[[y]], fun = function(x) x == 1)
    pps <- lapply(list(a, b), function(z) {
      ppp(x = z[, 1], y = z[, 2], window = sa_owin)
    })
    names(pps) <- c("gti", names(x)[y])
    reord <- sort(sapply(pps, function(i) i$n), decreasing = TRUE)
    ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
    mean(ppnn$dist) / 1000
  })
  names(mudist) <- names(x)[2:5]
  mudist
})  
names(mudists) <- c("yield", "prod", "carb", "eff")

@@

# output plot
cols <- c("red", "orange3", "green4", "blue")
pcol <- makeTransparent(cols, alpha = c(60))
plot(1:4, pch = 20, col = pcol)
bs <- do.call(rbind, mudists[c(1, 3)])
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")

pdf(full_path(p_fig, "location_offset.pdf"), height = 4, width = 4, 
    bg = "transparent")
par(mar = c(4, 4, 1, 0), bg = "transparent")
bp <- barplot(bs, beside = TRUE, col = ggplot2:::interleave(cols, pcol), las = 2,
              border = "transparent", xaxt = "n", ylab = "km", ylim = c(0, 16), 
              yaxt = "n")
axis(1, at = colMeans(bp), labels = lcnms, lty = 0)
axis(2, at = seq(0, 16, 2), labels = seq(0, 16, 2), las = 2)
legend(x = 9.5, y = 16, legend = rep("", 4), pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0, horiz = TRUE)
legend(x = 9.5, y = 15.25, legend = rep("", 4), pch = 15, adj = 0, col = pcol, 
       pt.cex = 1.5, bty = "n", cex = 0.8, horiz = TRUE, x.intersp = 0)
text(x = 10, y = 15.1, labels = "Yield", pos = 2, cex= 0.8)
text(x = 10, y = 14.35, labels = "Carbon", pos = 2, cex = 0.8)
dev.off()

>>>>

# yield and carbon data
load("external/ext_data/yieldprod-1km.rda")
load("external/ext_data/carbon-tier1.rda")
load("external/ext_data/cropland-1km.rda")
```

@@

### Identify high value areas
```{r, eval = FALSE}
# top decile of...
# ...cropland
crclass <- lapply(cropl, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

# ...yield
yclass <- lapply(cyld_agg$f1, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

@@

mudists <- lapply(list(crclass, yclass, pclass, cclass, effclass), function(x) {
  # x <- cclass
  mudist <- sapply(2:5, function(y) {  # y <- 2
    print(paste0("...", y))
    a <- rasterToPoints(x$gti, fun = function(x) x == 1)
    b <- rasterToPoints(x[[y]], fun = function(x) x == 1)
    pps <- lapply(list(a, b), function(z) {
      ppp(x = z[, 1], y = z[, 2], window = sa_owin)
    })
    names(pps) <- c("gti", names(x)[y])
    reord <- sort(sapply(pps, function(i) i$n), decreasing = TRUE)
    ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
    mean(ppnn$dist) / 1000
  })
  names(mudist) <- names(x)[2:5]
  mudist
})  
names(mudists) <- c("cropland", "yield", "prod", "carb", "eff")

@@

# output plot
cols <- c("red", "orange3", "green4", "blue")
pcol <- makeTransparent(cols, alpha = c(120))
pcol2 <- makeTransparent(cols, alpha = c(30))

plot(1:4, pch = 20, col = pcol)
bs <- do.call(rbind, mudists[c(1, 2, 4)])
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")

pdf(full_path(p_fig, "location_offset.pdf"), height = 3.5, width = 4, 
    bg = "transparent")
par(mar = c(2, 3.5, 1, 0), bg = "transparent")
bp <- barplot(bs, beside = TRUE, col = ggplot2:::interleave(cols, pcol, pcol2), 
              las = 2, border = "transparent", xaxt = "n", ylab = "", 
              ylim = c(0, 18), yaxt = "n")
axis(1, at = colMeans(bp), labels = lcnms, lty = 0, cex.axis = 0.9, 
     mgp = c(2, 0.25, 0))
axis(2, at = seq(0, 18, 2), labels = seq(0, 18, 2), las = 2)
mtext("km", side = 2, line = 2.25)
legend(x = 12.5, y = 18, legend = rep("", 4), pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0, horiz = TRUE)
legend(x = 12.5, y = 17.25, legend = rep("", 4), pch = 15, adj = 0, col = pcol, 
       pt.cex = 1.5, bty = "n", cex = 0.8, horiz = TRUE, x.intersp = 0)
legend(x = 12.5, y = 16.5, legend = rep("", 4), pch = 15, adj = 0, col = pcol2, 
       pt.cex = 1.5, bty = "n", cex = 0.8, horiz = TRUE, x.intersp = 0)
text(x = 13, y = 17, labels = "Cropland", pos = 2, cex= 0.8)
text(x = 13, y = 16.25, labels = "Yield", pos = 2, cex = 0.8)
text(x = 13, y = 15.5, labels = "Carbon", pos = 2, cex= 0.8)
dev.off()

### 2017-01-27 Rmd/c4-cropland-2.Rmd
Edits to manuscript to 27 Jan, 2017. Added/edited code in c3-cropland-1.Rmd (saved aggregated cropland and area weights raster lists to output rda files), c4-cropland-2.Rmd (calculated overall cropland areas by dataset and with different scales of aggregation).

# Load in datasets
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
load(full_path(p_data, "d_grid_abs.rda"))  # actual diffence grids
load(full_path(p_data, "bias_stats.rda"))  # bias statistics
load(full_path(p_data, "bias_stats_gti.rda"))  # bias statistics
# cellStats(!is.na(dlist_act$modmu$f1$g2011), sum)
```

@@

### Total error
Calculate how much cropland estimates differ between datasets
```{r, eval = FALSE}
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
gti <- lapply(gti, function(x) mask(x, namask))  # apply namask
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)

# recreate cropland percentages
lc2007 <- lapply(snms, function(x) gti$g2007 - dlist_act[[x]]$f1$g2007)
lc2011 <- lapply(snms, function(x) gti$g2011 - dlist_act[[x]]$f1$g2011)
# plot(lc2007[[2]] - lc2011[[2]])  # same thing, as it should be

# calculate km2 cropland in LC datasets
lckm2 <- sapply(lc2011, function(x) cellStats(x / 100, sum))
refkm2 <- sapply(gti, function(x) cellStats(x / 100, sum))
refkm2[2] / refkm2[1]  # 3.6 percent more in 2011 than 2007

# bias in total cropland relative to GTI
stats2007 <- round((refkm2[2] - lckm2) / refkm2[2] * 100, 1)
stats2011 <- round((refkm2[1] - lckm2) / refkm2[1] * 100, 1)

# N km^2 in country
#sapply(lc2007, function(x) cellStats(!is.na(x), sum))
# sapply(lc2011, function(x) cellStats(!is.na(x) , sum))
# sapply(gti, function(x) cellStats(!is.na(x) , sum))
totkm2 <- cellStats(!is.na(gti$g2011) , sum)
save(stats2007, stats2011, totkm2, refkm2, 
     file = full_path(p_data, "total-cropland.rda"))
```
<a href="#top">Back to top</a>

@@

```{r, eval = FALSE}
# Create weights maks
fact <- c(5, 10, 25, 50, 100)
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
# cellStats(!is.na(gti$g2011), sum)
# cellStats(namask, sum)

>>>>

# Load in datasets
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
load(full_path(p_data, "d_grid_abs.rda"))  # actual diffence grids
load(full_path(p_data, "bias_stats.rda"))  # bias statistics
load(full_path(p_data, "bias_stats_gti.rda"))  # bias statistics
load(full_path(p_data, "agg_cropland.rda"))  # aggregated cropland
# cellStats(!is.na(dlist_act$modmu$f1$g2011), sum)
```

@@

### Total error
Calculate how much cropland estimates differ between datasets
```{r, eval = FALSE}
# paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
# gti <- lapply(paths, raster)
# names(gti) <- c("g2007", "g2011")
# gti <- lapply(gti, function(x) raster::mask(x, namask))  # apply namask
# snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)

# recreate cropland percentages
# lc2007 <- lapply(snms, function(x) gti$g2007 - dlist_act[[x]]$f1$g2007)
# lc2011 <- lapply(snms, function(x) gti$g2011 - dlist_act[[x]]$f1$g2011)
# plot(lc2007[[2]] - lc2011[[2]])  # same thing, as it should be

# calculate km2 cropland in LC datasets
awgts <- lapply(area_wgts, function(x) x[[1]])
lckm2 <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
  gtia <- cellStats(gti_agg[[x]]$g2011 / 100 * awgts[[x]], sum)
  lca <- sapply(snms, function(y) {
    cellStats(lc_agg[[x]][[y]] / 100 * awgts[[x]], sum)
  })
  o <- c(gtia, lca)
}))

# lckm2 <- sapply(lc_agg$f1[snms], function(x) cellStats(x / 100, sum))
# lckm2 <- sapply(lc2011, function(x) cellStats(x / 100, sum))
# refkm2 <- sapply(gti, function(x) cellStats(x / 100, sum))

# area of each GTI dataset
refkm2 <- sapply(gti_agg$f1, function(x) cellStats(x / 100, sum))
refkm2[2] / refkm2[1]  # 3.7 percent more in 2011 than 2007

# bias in total cropland relative to GTI, at 1 km scale
stats2007 <- round((refkm2[2] - lckm2[1, ]) / refkm2[2] * 100, 1)
stats2011 <- round((refkm2[1] - lckm2[2, ]) / refkm2[1] * 100, 1)

# N km^2 in country
#sapply(lc2007, function(x) cellStats(!is.na(x), sum))
# sapply(lc2011, function(x) cellStats(!is.na(x) , sum))
# sapply(gti, function(x) cellStats(!is.na(x) , sum))

totkm2 <- cellStats(!is.na(gti_agg$f1$g2011) , sum)
save(stats2007, stats2011, totkm2, refkm2, 
     file = full_path(p_data, "total-cropland.rda"))
```
<a href="#top">Back to top</a>

@@

```{r, eval = FALSE}
# Create weights mask
fact <- c(5, 10, 25, 50, 100)
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
# cellStats(!is.na(gti$g2011), sum)
# cellStats(namask, sum)
